\documentclass[12pt]{article}
\usepackage[centertags]{amsmath}
\usepackage{amsfonts,amssymb,amsthm,color}

%
%MY SHORTCUTS
%
\newcommand{\mcE}{\mathcal{E}}
\newcommand{\mcC}{\mathcal{C}}
\newcommand{\mcA}{\mathcal{A}}
\newcommand{\mcB}{\mathcal{B}}
\newcommand{\mcM}{\mathcal{M}}
\newcommand{\mcN}{\mathcal{N}}
\newcommand{\mcL}{\mathcal{L}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\satisfies}{\models}
\newcommand{\A}{\forall}
\newcommand{\E}{\exists}
\newcommand{\isom}{\cong}
\newcommand{\diag}{\text{Diag}}
\newcommand{\Th}{\text{Th}}
\newcommand{\eldiag}{\text{Diag}_\text{el}}
\newcommand{\qe}{quantifier elimination }
\newcommand{\qf}{quantifier-free }
\newcommand{\omst}{\(\omega\)-stable }
\newcommand{\acl}{\text{acl}}

%
%PAGE REFORMATTING
%
\setlength{\topmargin}{0 in}
\setlength{\headheight}{0.25 in}
\setlength{\oddsidemargin}{0.5 in}
\setlength{\textwidth}{6 in}
\setlength{\textheight}{8.5 in}
\setlength{\headsep}{0.25 in}
\newcommand{\singlespaced}{\renewcommand{\baselinestretch}{1}\normalfont}
\newcommand{\doublespaced}{\renewcommand{\baselinestretch}{2}\normalfont}
\singlespaced
\doublespaced

%
%THEOREM STYLES
%
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{question}[theorem]{Question}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{example}[theorem]{Example}
\newtheorem{convention}[theorem]{Convention}

\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{fig}{Figure}[section]
%
\theoremstyle{plain}
\newtheorem{res}[defn]{Result}
\newtheorem{thm}[defn]{Theorem}
\newtheorem{cor}[defn]{Corollary}
\newtheorem{prop}[defn]{Proposition}
\newtheorem{lem}[defn]{Lemma}
%
\theoremstyle{remark}
\newtheorem*{case}{Case}
\newtheorem{rem}[defn]{Remark}
\newtheorem{ex}[defn]{Example}
\newtheorem{exs}[defn]{Examples}

%
%EQ. NUMBERING
%
\numberwithin{equation}{section}
\renewcommand{\theequation}{\thesection.\arabic{equation}}
%
\def\thetitle{A Proof of Morley's Categoricity Theorem}
\def\theauthor{Adam Freilich}
\def\theadvisor{Advisor: Dr. Henry Towsner}
\def\theyear{2014}
%
\pagenumbering{roman}

\begin{document}
\include{Titlepage}
%\include{Abstract}
%\include{Dedication}
\pagenumbering{arabic}
\pagestyle{myheadings} \markright{}

\section{Basic Notions}
\input{chapters/introduction}

\section{\(\omega\)-stability}
\input{chapters/stability}

\section{Vaughtian Pairs}
\input{chapters/vaughtian-pairs}

\section{Proof of Morley's Categoricity Theorem}
\input{chapters/proof-morley}

\section{Further Questions}
\input{chapters/extensions}

\appendix
\section{Appendix}
\input{chapters/appendix}

\begin{thebibliography} \small
%
\bibitem[Mar]{mar} David Marker, Model Theory: An Introduction, 2002

\bibitem[Sh1]{shelahUncountable} Shelah, Saharon. Categoricity of uncountable theories -- Proc Tarski Symposium (Univ. of California, Berkeley, Calif., 1971) (1974) 187--203

\bibitem[Sh2]{shelahUnsuperstable} Shelah, Saharon. "Why there are many nonisomorphic models for unsuperstable theories." Proceedings of the International Congress of Mathematicians, Vancouver. 1974.

\bibitem[Har]{hart}Hart, Bradd. A Proof of Morley's Conjecture. Journal of Symbolic Logic 54 (1989), no. 4, 1346--1358. http://projecteuclid.org/euclid.jsl/1183743103.
%
\end{thebibliography}
\end{document}

